\chapter{Deliverables}
The main deliverable is a working spatial channel sounding prototype, already in the Radio Science Lab. It consists of the following components.
\section{Hardware}
\subsection{Transmitter}
   \begin{itemize}
   \itemsep-1em 
      \item Mini-PC
      \item USRP
      \item Rubidium Frequency Standard
      \item GPS Receiver
      \item UPS
      \item Antenna
      \item Antenna Mast
      \item Amplifier
   \end{itemize}
\subsection{Receiver}
   \begin{itemize}
   \itemsep-1em 
      \item Mini-PC
      \item USRP
      \item Rubidium Frequency Standard
      \item GPS Receiver
      \item UPS
      \item Antenna
      \item Antenna Mount for Dr. Michelson's car
   \end{itemize}
\section{Software}
   \subsection{Transmitter}
      \begin{itemize}
      \itemsep-1em 
         \item \texttt{transmitter\_matlab.m}, Transmitter signal senerator (MATLAB)
         \item \texttt{transmitter.sln}, USRP transmitter software (C++)
      \end{itemize}
   \subsection{Receiver}
      \begin{itemize}
      \itemsep-1em 
         \item \texttt{verifySignal.m}, for verifying signal validity (MATLAB)
         \item \texttt{receiver.sln}, USRP receiver (C++)
      \end{itemize}
   \subsection{Post-Processing and Image Reconstruction}
      \begin{itemize}
      \itemsep-1em 
         \item \texttt{prepSignalForSim.m}, for range-compressing received signals (MATLAB)
         \item \texttt{jeppler\_hsrsar\_backproj.m}, Simulation and image reconstruction algorithm received from MDA.
         \item \texttt{backscatter.sln}, Fast image reconstruction algorithm re-written in C++ using DirectX and GPU.
      \end{itemize}
\section{Documentation}
   \begin{itemize}
   \itemsep-1em
      \item GPU programming concepts and documentation of our reconstruction algorithm.
      \item Proposal
      \item Final Report
   \end{itemize}
\section{Estimated Raw Cost}
Since all of these components were already available in the RSL at the time of project commencement, there are no current project costs. Table \ref{tab:total_cost} shows the estimated total cost of all parts for both the transmitter and receiver.

\begin{table}[h!]
\begin{center}
\begin{tabular}{l|l}
Item & Cost \\ \hline
2x Lenovo M93p & 2x \$800 \\
2x USRP N210 & 2x \$1700 \\
2x Rubidium Frequency Standard & 2x \$1500 \\
2x UPS & 2x \$100 \\
2x Timing GPS & 2x \$200\\
2x Panel Antenna & 2x \$200 \\
30W Amplifier & \$3000 \\
DRGPS & \$400 \\\hline
Total Cost & \$12400
\end{tabular}
\end{center}
\caption{\label{tab:total_cost} Estimated total cost of prototype system}
\end{table}
